---
title: 'Testing Rmarkdown with GNIP data'
date: '2021-02-20'
---

```{r, setup, echo=FALSE, include=FALSE}

knitr::opts_chunk$set(echo=TRUE)
pacman::p_load(here, knitr, blogdown, tidyverse, kableExtra, ggthemes)
```

## Viewing data

I've downloaded a small set of data from the [International Atomic Energy Agency Global Network for Isotopes in Precipitation](https://nucleus.iaea.org/wiser) that includes water isotope data collected in the last 10 years from 14 sites around the world. With this data source, I can't pull directly, but I can convert and load a .csv. The code chunk that produces this table also cleans up the data for display. The GNIP source doesn't specify units (!!) so I have had to guess. 

```{r GNIP-all, echo=FALSE}
waters.all <- read.csv(here("data/water/GNIP-daily_2010-2020.csv"))

kable(waters.all %>%
      select(Site, Country, Date, 
             "Altitude (m)" = Altitude, 
             "d18O VSMOW" = O18, 
             "dD VSMOW" = H2, 
             "Precipitation (mm)" = Precipitation, 
             "Air Temperature (C)" = Air.Temperature, 
             "Vapor pressure" = Vapour.Pressure),
      digits = 2,
      format="html") %>%
      kable_styling() %>%
      scroll_box(height="500px", width="100%") 
```

Getting this to look better will require adjusting the table style in the theme, and messing with kable or another package. Having special characters in a table display would be nice, but since they aren't allowed in column names for raw data, it may not be the best precedent to set. 

## Post plot 

A basic ggplot for proof of concept.

```{r basic-plot, echo=FALSE}
theme_set(theme_tufte(base_family = "Helvetica"))

axis.x.max <- max(waters.all$O18) + 0.5
axis.x.min <- min(waters.all$O18) - 0.5
axis.y.max <- max(waters.all$H2) + 0.5
axis.y.min <- min(waters.all$H2) - 0.5

axis.x.label <- expression(paste(delta^{18}, "O (‰ VSMOW)"))
axis.y.label <- expression(paste(delta, "D (‰ VSMOW)"))

plot.basic <- ggplot(data=waters.all, aes(x=O18, y=H2, color=Country)) +
  geom_point() +
  geom_abline(data=NULL, aes(slope=8, intercept=10, color="GMWL")) +
  scale_color_colorblind() +
  
  # axes and titles
  coord_cartesian(xlim=c(axis.x.min, axis.x.max), ylim=c(axis.y.min, axis.y.max)) + 
  theme(legend.title=element_blank(), 
        panel.border=element_rect(fill = NA)) + 
  labs(x = axis.x.label, 
       y = axis.y.label)

plot(plot.basic)

```

OK! 